---
title: "week_05"
author: "Venkata Santosh Kumar Kornu"
format: html
---

## Overview {.unnumbered}

### What this file is for

This week is hands-on with **Bias**,**Fairness**,**Ethics** **&** **Resproducibility**. We have Did:

-   Detect imbalance vs. bias and relate it to project purpose.
-   Explain fairness notions (representation vs outcome fairness).
-   Perform hypothesis tests and interpret confidence intervals.
-   Use a fixed random seed to make analyses reproducible.
-   Observe how different seeds affect results and train/test splits.
-   Save a lightweight reproducibility log (session info) to your repo.

```{r}
library(tidyverse)
library(dbplyr)
data <- read_csv("data/processed/cleaned_data.csv")
glimpse(data)
```

```{r}
library(tidyr)

data_long <- data |>
  pivot_longer(
    cols = -Year,
    names_to = "Country",
    values_to = "Emission"
  )
```

## Imbalance vs Bias

## Descriptive distribution

```{r}
tbl <- table(data_long$Country)
barplot(tbl, main = "Group Distribution (Country)", ylab = "Count",
col = "lightblue", border = "white")
round(prop.table(tbl), 2)
```

The file have equal or balanced number of values among all country and equal amount of data collected in all country. There is no difference among countries. There is no **Bias** in the data.

## Fairness - Random Split + Seed

```{r}

set.seed(10) # ensures your peers obtain the same random split
data_cat <- data |> mutate(group = sample(c("A", "B"), n(), replace = TRUE))
data_cat |>
group_by(group) |>
summarise(avg_Germany = mean(Germany, na.rm = TRUE))
```

we have created or splitted the data into two group that is Group A and Group B

Group A is 71273452 Group B is 84076017

```{r}
boxplot(Germany ~ group, data = data_cat,
        main = "Germany Emissions by Group",
        xlab = "Group",
        ylab = "Emission",
        col = c("lightblue", "lightgreen"))

```

This graph showing boxplot of to different groups.

## Reproducibility Log

```{r}
set.seed(42)
sample_data <- dplyr::sample_n(data_long, 30)
# (2) Simple model
model <- lm(Emission ~ Year, data = sample_data)
summary(model)$coefficients

# (3) Save reproducibility details (seed + session info)
sink("session-info.txt")
cat("=== Reproducibility Log ===\n")
cat("Random seed used: 42\n")
cat("Date:", Sys.Date(), "\n\n")
sessionInfo()
sink()
# Confirm file creation (returns TRUE if created)
file.exists("session-info.txt")
```

The Intercept is 6619647225 that means emission happen at 0 year.It just helps define the regression line.

The Slop is -3271592 This means decrease by about 3.27 million units per year on average.

t value is -1.197 that means how strongly year explains Emission change.

## Hypothesis Testing & Confidence Intervals

```{r}
cor.test(data$Year, data$Germany)

```

The Pearson's correlation between Year and Germany in your dataset. cor = -0.0533, this says very weak negative correlation between year and Germany where year increase the germany values tends to decrease.

```{r}
t.test(Emission ~ Country, data = subset(data_long, Country %in% c("Germany", "Belgium")))

```

The Weich t-test compares the Emission betwween Belgium and Germany. The p-value (**4.206e-07**) which is very small than 0.05 this means the difference in emission between Belgium and Germany is **highly statiscally** significant.

The mean of both Belgium (9071058) and Germany (77333333) says Germany emits far more than Belgium.

```{r}
set.seed(10)
m1 <- lm(Emission ~ Year, data = dplyr::sample_n(data_long, 30))

set.seed(99)
m2 <- lm(Emission ~ Year, data = dplyr::sample_n(data_long, 30))

summary(m1)$coefficients
summary(m2)$coefficients
```

# Model 1 

Intercept is 1692630790.5

Year is -836756 this mean for every increase in unit in year , the dependent variable decreases by 836756 units.

p-value is 0.3098 > 0.05, so not statistically significant.

# Model 2

Intercept is -1006991705.2

Year is +501903 this mean for every increase in unit in year , the dependent variable increase by 501903 units.

p-value is 0.5159 > 0.05, so not statistically significant.

From both **Model 1** is slightly decreased per year and **Model 2** is slightly increase per year.

## Conclusion

-   Changing the random seed changes which subset of data gets sampled, slightly altering the model. Fixing a seed (e.g., set.seed(42)) makes results reproducible.
-   Equal data distribution across countries supports fair analysis.
-   Simple linear regression indicates minor year-to-year variation in emissions
