---
metadata-files: ["_metadata.yml"]
title: "Week 2 – Dataset Exploration and First Wrangling"
subtitle: "Selecting, Cleaning, and Understanding Your Data"
author: "Venkata Santosh Kumar Kornu"
---

# Overview
*Introduce the purpose of Week 2 and what this document covers.*

- Key aims of this week  
- Expected outcomes  
- General encouragement statement

---

# Dataset Selection and Rationale
## 1.1 Dataset choice
*Explain what dataset you are using and why.*
the dataset I have choose is from Eurostat - Greenhouse Gas Emission from transport,energy and agriculture and other industries . I took this dataset because it is important and more focusing topic which is essential for improving nature.

### (Insert code cells – load dataset)
```{r}
# Example: load dataset
library(readxl)
data <- read_excel("data/raw/data_copy.xlsx")

```


```{r}
      
library(dplyr)
str(data)
nrow(data)
ncol(data)
```

## 1.2 Justification

Describe why this dataset suits your project idea.

this data has everthing I want to do anaylsis like 
it consist of 66 sheets which covers all three different type of emission like Co2, Methane, Nitrogenous oxide by different indrustries.

# 2 First Glimpse of the Data
## 2.1 Initial inspection

Discuss the structure, dimensions, and variable types.

(Insert code cells – glimpse, summary)
```{r}
# Example: check structure

summary(data)


```

```{r}
glimpse(data)
```


```{r}
head(data, 10)
```

## 2.2 Observations

Summarise key insights about shape, columns, and data types.
The dataset consist of 33 column and 150 row by taking 15 sheets data and converting into one sheet
data types is numeric and character type.

# 3 Initial Data Wrangling

Document cleaning, transformations, and reasoning.

## 3.1 Cleaning

Describe how you handled missing values, duplicates, formats.

In my data,at first it didn't show me missing values because the character present in valuse ":", and then coverting ":" into NA , I found 45 missing values.

and I used imputation method by using mean of respective features

I checked for duplication and results return "FLASE" that mean no duplicates

(Insert code cells – cleaning steps)

```{r}

data[data == ":"] <- NA
colSums(is.na(data))




```

```{r}
data_clean <- na.omit(data)

year_cols <- 2:33   # adjust if your columns are different
data[, year_cols] <- lapply(data[, year_cols], function(x) as.numeric(as.character(x)))
```



```{r}
data$Turkiye <-
  ifelse(is.na(data$Turkiye) ,  mean(data$Turkiye, na.rm = TRUE),data$Turkiye)

data$Serbia <-
  ifelse(is.na(data$Serbia) ,  mean(data$Serbia, na.rm = TRUE),data$Serbia)

data$Switzerland <-
  ifelse(is.na(data$Switzerland ) ,  mean(data$Switzerland , na.rm = TRUE),data$Switzerland)


```


```{r}
data[data == ":"] <- NA
colSums(is.na(data))    ##checking is there missing values
```



```{r}
duplicated_row <- duplicated(data)

any(duplicated_row)

data[duplicated_row,]
```



## 3.2 Transformation, integration, validation

Conceptual or implemented; note decisions or possible future steps.

(Insert code cells – optional transformation)

```{r}
# derived variables or merges

```

## 3.3 Storing cleaned data

Describe where you saved the processed file and why.
the processed file is saved in processed folder in the data folder as per the insturtion.

(Insert code cells – write processed file)

```{r}
# write cleaned dataset

library(readr)
data_joined <- data
write_excel_csv(data_joined, "data/processed/cleaned_data.csv")
```

# 4 Exploratory Data Analysis (EDA)
## 4.1 Variable identification

List variable types and their roles.

(Insert code cells – variable listing)
```{r}
# identify variable types
glimpse(data)
```

## 4.2 Descriptive statistics

Discuss central tendencies, dispersion, outliers.

(Insert code cells – summarise numeric variables)
```{r}
# descriptive statistics
summary(data)
```
## 4.3 Numerical analysis

Discuss histograms and boxplots

```{r}
hist(
  data$Germany,
  main = "Total Emission ",
  xlab = " Emission (years)"
)
```
```{r}
library(tidyr)
library(dplyr)

data_long <- data %>%
  pivot_longer(
    cols = -Year,
    names_to = "Country",
    values_to = "Emission"
  )

```



```{r}
boxplot(
  Emission ~ Country,
  data = data_long,
  main = "Emission distribution by Country",
  xlab = "Country", ylab = "Emission",
  las = 2
)
```


(Insert code cells - histograms and baxplots)

## 4.4 Categorical analysis

Discuss frequency distributions or group counts.

(Insert code cells – counts or plots)

```{r}
# bar plot / counts
data_summary <- data_long %>%
  group_by(Country) %>%
  summarise(mean_emission = mean(Emission, na.rm = TRUE))

barplot(data_summary$mean_emission,
        names.arg = data_summary$Country,
        main = "Average Emission by Country",
        ylab = "Mean Emission",
        las = 2)


```

## 4.5 Interpretation

Write short commentary linking patterns to your project context.

# 5 Documentation and Reflection

Capture reasoning, challenges, and decisions.

## 5.1 What assumptions did you make?
Assumptions are made in identifying the missing values after showing zero missing values after deep research got to know that values column consists some char ":" and replaced it with mean.

## 5.2 What was easy or hard?
loding data into r server is easy and using summary and glimpse made task easier to understand the data like data type , mean , medium, max and min. the hard part is formatting data , i have 66 sheets in that 21 sheets are CO2 which is taken and converted 15 sheets into one sheet.
plots part is difficuilt to get info so couldn't able to do.

## 5.3 What would you do differently next time?
Next time,I will understand how plots should be done and get better understandable and meaningful analysis
and will also add few more data sheets of different emission like methane etc


# 6 Good Practice Reminders

Checklist and reproducibility notes.

* Comments in code 
* Relative paths
* Push and pull discipline
* Consistent naming conventions

# 7 Group Notes and Variations

Describe context-specific adaptations or creative extensions.


# 8 Final Checkpoint Before Submission

Tick items off to ensure readiness.

* Dataset in /data/raw/  YES
* Cleaned dataset in /data/processed/ YES
* HTML output in /outputs/ YES
* Reflection written/  yes
* Changes committed and pushed/ Yes

