---
title: "Week 4 – Statistics: Distributions & First Model"
author: "ods-team-48 / Venkata Santosh Kumar Kornu/AKHILESWARA REDDY LINGIREDDY/SOURABH SHRINIVAS KULKARNI/"
date: "`r format(Sys.Date())`"
format:
  html:
    toc: false
    toc-depth: 2
---

## Overview {.unnumbered}

::: callout-note
### What this file is for



This week is hands-on with **statistics**. You will:
- Measure spread (**SD, IQR**) for two numerical variables
- Visualise **distribution shapes** (histograms + Normal overlays)
- Fit a **simple linear regression** (e.g., `grade ~ study_hours`)
- Inspect **residuals** to check model fit
- Record brief **group reflections**

Work with your **cleaned Week 2 dataset** (no impossible or extreme values).
:::

---

# Dataset & Setup

Use your cleaned dataset prepared in Week 2. Mind the file path.

```{r}
library(tidyverse)

# Load your cleaned dataset

data = read_csv("data/processed/cleaned_data.csv")
# Quick structure check
glimpse(data)
```

```{r}
head(data)
```


**Confirm (1–2 lines):** that the dataset loaded correctly (rows/columns) and that `grade` and `study_hours` are numeric.

the dataset is loaded correctly , rows and columns are cleaned .

---

# Spread & Variability

Compute summaries, SD, and IQRs for at least two two main variables from your dataset (e.g., `grade` and `study_hours`).

After computing I got SD of Germany is 157047697 IQR of Germany is 69106277

```{r}
# Summaries
summary(data$Germany)
sd(data$Germany, na.rm = TRUE)
IQR(data$Germany, na.rm = TRUE)

summary(data$Year)
sd(data$Year, na.rm = TRUE)
IQR(data$Year, na.rm = TRUE)


```

**Comment (2–3 lines):** Which variable varies more? Are the ranges reasonable? What does the **IQR** say about the middle 50%?

The variable Germany varies far more than Year

The range of Germany is 666026516 which is extremely large.

IQR of Germany is 69106277 which says that middle 50% of emission values are highly spread out.
---

```{r}
boxplot(data$Germany)
```

# Distribution Shapes

Plot histograms and compare with a Normal curve for the same variables.

```{r}
# Histogram and normal curve for variable 1
hist(data$Germany, breaks = 15, prob = TRUE,
main = "Total Emission by Germany", xlab = "Germany", col = "skyblue",
border = "white")
curve(dnorm(x, mean(data$Germany), sd(data$Germany)),
add = TRUE, col = "red", lwd = 2)

```

```{r}
Q1 <- quantile(data$Germany, 0.25)
Q3 <- quantile(data$Germany,0.75)
IQR <- Q3 - Q1

lower_bound <- Q1 - 1.5* IQR
upper_bound <- Q3 + 1.5 * IQR

data_no_outliers <- subset(data,Germany >= lower_bound & Germany <= upper_bound)
```



**Comment (2–3 lines):** Is each variable roughly Normal, skewed, or heavy-tailed? Does the red curve fit well? Any gaps or outliers?

The graph show the Right skewed, there are gaps and outliers in the graph.red curve doesn’t fit well.
---

```{r}
boxplot(data_no_outliers$Germany)
```

```{r}
cor(data)
```

# First Model (Simple Linear Regression)

Fit a simple linear model (use appropriate variables and reason your choice).

```{r}
# create a model using lm and provide summary
model <- lm(Germany ~ Year, data = data)
# Inspect key results
summary(model)


```

```{r}
model <- lm(Germany ~ Year, data = data_no_outliers)
# Inspect key results
summary(model)
```

```{r}
plot(data$Year, data$Germany,
xlab = "Total Emission per Year", ylab = "Germany",
main = "Yearly Emission vs Germany", pch = 19, col = "darkgrey")
abline(model, col = "red", lwd = 2)
```

```{r}
plot(data_no_outliers$Year, data_no_outliers$Germany,
xlab = "Total Emission per Year", ylab = "Germany",
main = "Yearly Emission vs Germany", pch = 19, col = "darkgrey")
abline(model, col = "red", lwd = 2)
```
```{r}
coef(model)
```


**Interpretation (3–4 lines):**
Model with outliers:

- Intercept (β₀) = 5943752819 means that when Year = 0, the predicted emission value for Germany would be about 5.94 billion.

- Slope (β₁) = –2.91 × 10⁶ means for each passing year, Germany’s emissions are predicted to decrease by about 2.9 million units on average.

- (Optional) Is the slope statistically significant (p-value)? the p- value is 0.517 is greater than 0.05 so the slope is not statistically significant, meaning there’s no strong evidence of a linear trend

- (Optional) How much variation does the model explain (R²)? The model give R^2 value as 0.0028

Model without Outliers:

- Intercept (β₀) = -2.301 × 10⁹ means when Year = 0, the predicted emission value for Germany would be about –2.3 billion,

- Slope (β₁) = 1.147 × 10⁶ means every year, Germany’s emissions are predicted to increase by about 1.15 million units on average.

- (Optional) Is the slope statistically significant (p-value)? the p-value is 0.202

- (Optional) How much variation does the model explain (R²)? the model R^2 is 0.0135

After removing outliers, the trend is slightly upwards.

::: callout-note
### What is a model?

A **model** is a simplified, mathematical explanation of how things relate or behave.  
It is **expressed mathematically** (equation/algorithm), **summarises** relationships, **simplifies** what matters, **generalises (within limits)**, and **must be interpretable**.

> A good model is clear, honest, and useful — not a perfect copy of reality, but a way to think better about it.
:::

---

# Model Fit (Residuals)

Residuals are **errors** = observed − predicted. They should look like random noise around 0.

```{r}
# Plot the residuals including abline
plot(model$residuals/1e6,
main = "Residuals from Regression Model",
ylab = "Residual (error)",
xlab = "Index (observation)",
col = "purple", pch = 19)
abline(h = 0, col = "red", lty = 2)

```

**Comment (2–3 lines):** Are residuals roughly centred around 0 and randomly scattered? Any patterns or extreme points? What does this suggest about model fit?

The residuals is not centered around 0, There is two extreme positive residual at 150 which is clear outlier that means residuals are not evenly scattered.
---

# Group Reflection

Keep answers short and specific (1–2 sentences each).

| Question | Reflection |
|:--|:--|
| **Distributions:** What did shapes reveal about variation or consistency? | |

|Boxplot shows there are so many outliers in the Germany’s data values and after removing there outliers, the distribution became more consistent and withh less spread and fewer extreme values.|

| **Model:** What does your regression line suggest about the relationship? | 

|The regression line with outlier and without outlier have a different results like the line before outliers removed showed very weak negative trend while after cleaning, it indicated a slight positive trend. But in both case, the relationship between year and emission was not statistically significant.||

| **Learning:** What was the main thing you learned about your dataset this week? | |

|I have learned that outliers can heavily influence regression results and distort trends.Cleaning the data improves interpretation but doesn’t always create a significant relationship|

---

# Expected Outputs

Your file should include (at least):
- Two distribution plots
- One scatterplot with fitted regression line
- One residual plot
- Regression summary with brief interpretation
- Completed reflection table

---

# You are free to explore further along modeling (e.g., trying different variable combinations, or trying multiple regression, or different models available in R libraries).
# You may also explore along p-values and R².
