
---
title: "Week 7 – Visualisation & Storytelling"
author: "Venkata Santosh Kumar Kornu"
format:
  html:
    toc-depth: 2
    code-fold: false
    code-tools: false
execute:
  echo: true
  warning: false
  message: false
---



*(Individual Submission)*

This file contains the required structure for your Week 7 submission.
Replace all placeholder text with your own work.

---

# Dataset Used

::: callout-important
You are expected to use the same dataset that you used during previous weeks. However, you may switch to another dataset if you can justify the switch. The new dataset (if switched to) must adhere to the same theme as before.
:::

Briefly describe the dataset you are visualising in the callout below (2–4 sentences). If switching datasets, complete the note below.

::: callout-note
### Dataset Description and Link
I am using the same dataset.

:::

::: callout-warning
#### Dataset Switch Note (Only if applicable)

* Why the original dataset could not be used responsibly
* Why this new dataset fits the same theme
* Source + license

:::

---

## Improved Visuals from Previous Weeks

*(Minimum 1, Maximum 3)*

Please provide reference to the visuals you choose to improve from previous submissions with traceable and clear reference.

::: {.callout-important collapse="true"}
#### What “improvement” of a previous plot means (click to reveal)

You may re-include 1 to 3 visuals from Week 2 or Week 4, but only if you improve them. In Week 7, “improvement” does not mean changing the data or re-running the same code. It means making the visualisation clearer, fairer, more accurate, or more purposeful.

You may improve a plot by:

- Fixing misleading or confusing scales
(e.g., avoiding truncated axes, using appropriate limits).

- Choosing a more suitable chart type
(e.g., boxplot instead of bar chart for comparisons).

- Improving aesthetics
(cleaner themes, readable labels, consistent font sizes).

- Clarifying colour choices
(e.g., use of high-contrast or colour-blind-friendly palettes).

- Adding informative titles or subtitles
(summarising the key message).

- Fixing labelling issues
(rename ambiguous variables, use meaningful captions).

- Removing clutter or chart junk
(superfluous gridlines, 3D effects, unnecessary backgrounds).

- Highlighting or annotating important features
(e.g., a noticeable outlier, a trend, or a threshold).

- Ensuring ethical clarity
(e.g., appropriate grouping, non-stereotypical colour use).

Your task is simply:

Take a plot you created earlier and improve how clearly, honestly, and effectively it communicates its insight. You may keep the same chart type or change it to a better one — explain your choice in the bullet points that follow.

:::
---
```{r}
library(readr)
library(tidyr)
library(ggplot2)
library(dplyr)
library(scales)
```


```{r}
data <- read.csv("data/cleaned_data.csv")
glimpse(data)

```

```{r}
summary(data)
```

```{r}
data_long <- data |>
  pivot_longer(
    cols = -Year,
    names_to = "Country",
    values_to = "Emission"
  )
data_long <- data_long |>
  
  mutate(
    Emission = as.numeric(as.character(Emission)),
    Year = as.integer(Year)  
  )
```

### 2.x Original Plot 1 (from Week 2 or 4)

```{r}
# Paste your original code here
 boxplot(
 Emission ~ Country,
 data = data_long,
 main = "Emission distribution by Country",
 xlab = "Country", ylab = "Emission",
 las = 2
 )

```

> *Include a short reminder of what the original plot attempted to show.*

 The plot attempted to show distribution of emission values for each country.
---

### 2.x Improved Plot 1 (Week 7)

```{r}
# Your improved plot code goes here

ggplot(data_long, aes(x = Country, y = Emission,fill = Country)) +
  geom_boxplot() +
  labs(
    title = "Distribution of Country Values by Year",
    x = "Country",
    y = "Emission"
  ) +
  theme_minimal()+
  theme(axis.text = element_text(angle = 45,hjust = 1))

```
### 2.x Original Plot 2 (from Week 2 or 4)

```{r}
 data_summary <- data_long %>%
 group_by(Country) %>%
 summarise(mean_emission = mean(Emission, na.rm = TRUE))
 barplot(data_summary$mean_emission,
 names.arg = data_summary$Country,
 main = "Average Emission by Country",
 ylab = "Mean Emission",
 las = 2)
```


### 2.x Improved Plot 2 (Week 7)

```{r}

ggplot(data_long, aes(x = Country, y = Emission,fill = Country)) +
  stat_summary(fun = "mean", geom = "bar")+
  geom_col()+
  labs(
    title = "Frequency of Emission done by country",
    x = "Country",
    y = "Emission"
  ) +
  theme_minimal()+
  theme(axis.text = element_text(angle = 90,hjust = 1))
```

**Include require text/code sections if you are adding more improved visuals from previous weeks.**
```{r}
data_rank <- data_long |>
  group_by(Country) |>
  summarise(TotalEmission = sum(Emission, na.rm = TRUE)) |>
  arrange(TotalEmission)

ggplot(data_rank, aes(x = reorder(Country, TotalEmission), 
                        y = TotalEmission, 
                        fill = TotalEmission)) +
  geom_col() +
  labs(
    title = "Countries Ranked by Total Emission (Low to High)",
    x = "Country",
    y = "Total Emission"
  ) +
  scale_fill_gradient(low = "green", high = "red") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


::: callout-note
#### What I Improved and Why (3–6 bullet points):

* I improved the labels of the chat to more readable
* Adjusted scaling in the graph.
* Added different colour shade to each country.
* Now created two bar graph one added different colors to the bars based on country and another created bar based on total emission by countries and arranged in low to high with specified colour like low is **Green** and high is **Red**
* So from above there are four countries which emits more they are **Germany** , **Turkiye**, **Poland**, **Italy** they emit more than other and **Germany** is in **danger zone**.

:::


---

**Include require text/code sections if you are adding more improved visuals from previous weeks.**

---

## New Week 7 Visuals

You must produce **at least two new required visuals**:

::: {.callout-important collapse="true"}

#### Possible New Plot Types for Week 7 (Click to reveal)

**(Choose the ones that fit your data best)**

Depending on the nature of your dataset (categorical, numerical, time-based, grouped, or mixed), you may choose from the following plot types:

**Relationship Visuals (for Required Plot A)**

- Scatterplot
- Scatterplot with regression line (lm)
- Scatterplot with LOESS smoothing (with visible span parameter)
- Scatterplot faceted by group
- Bubble chart (if one variable can act as size)
- Coloured scatterplot by category
- Pairwise scatterplot
- Hexbin or 2D density (for large datasets)

**Distribution Comparison (for Required Plot B)**

- Boxplot
- Violin plot
- Half-violin + boxplot hybrid
- Density overlays
- Ridge plot (if you have multiple groups)
- Histogram comparison (but avoid overlapping clutter)
- Dotplot / beeswarm (for small samples)

**You may add one optional visual, provided it is meaningful for your dataset:**

- Time series line chart
- Time series with responsible smoothing
- Grouped line chart
- Bar chart of counts (not means)
- Stacked or filled bar proportions
- Tile/heatmap (if you have two categorical variables)
- Treemap (for hierarchical counts)
- Scatterplot with model diagnostics (residuals, fitted vs. observed)
- Small multiples (facet grid)
- Correlation heatmap (only if your dataset has ≥4 numeric variables)

**If none of the above fits your dataset, you can choose another plot type — as long as:**

- it is appropriate for what you want to show,
- it is honest and clear,
- you justify why it fits your data.

:::

**Note:** If required plots are not possible, please justify the alternatives and adjust the text and explanations accordingly.

---
```{r}
unique(data_long$Country)
```

### A. Relationship Visual (Required)

> Examples: scatterplot + trend line, scatter by subgroup, faceted scatter.

```{r}
# Relationship plot code here

ggplot(data,aes(x = France,y = Italy))+
  geom_point()+
  geom_smooth(method = "lm",se = FALSE, linewidth = 1)+
  labs(
    title = "Relationship btw France and Italy",
    x = "France Values",
    y = "Italy Values" 
  )+
  theme_minimal()
```

**Interpretation (2–4 sentences):**
- The scatterplot shows how france emission relate to italy's emission across all years.
- each point represents each year and shows overall upwards trend tells when france emission increase then italy emission tends to increase.
- The line says a positive linear relationship between them.

**Why this plot type is appropriate (1–2 sentences):**
- Scatter plot is appropriate because it directly visualizes the relation between two numerical variables and with regression line helps reveal the overall trend.
---

### B. Distribution Comparison (Required)

> Choose one: boxplot, violin, density comparison, faceting.

```{r}
ggplot(data_long, aes(x = factor(Year), y= Emission,fill = factor(Year)))+
  geom_violin(trim = FALSE, alpha= 0.7)+
  labs(
    title = "Distribution Of Emission by Year using Violin",
    x = "Year",
    y = "Emission"
  )+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  coord_flip()+
  guides(fill= FALSE)

```
**Interpretation (2–4 sentences):**
- The Violin plot shows how emission levels vary across different years.by showing both spread and density of data.
- wider section in violin indicate year where emission values are more like in graph **2015** have wider section tells more emission.

**Why this plot type is appropriate (1–2 sentences):**

Violin plot is appropriate because it combines the advantage of boxplot with density curve which allows to compare both summary statistics and distribution shapes.
---

### C. Trend / Group Comparison (Optional)

```{r}
# Optional extra plot
```

> Explain only if included.

---

## Narrative

Write a short narrative connecting your visuals into a coherent data story.

Example structure (keep it simple):

* What question/theme the visuals address
- The visuals address country-level emission from 2014 to 2023.

* What the relationship plot reveals
- The plot shows a positive correlation between France and Italy emission ,that tell if emission rise in one country then other rises too.

* What the distribution plot clarifies
- Used violin plot shows how emission levels vary across different years.by showing both spread and density of data.

* Optional: what trend/group plot adds
* A closing insight
- This gives clear understanding of both absolute values and relative trends.
---

## Responsible Design Notes

Provide brief bullet points about your design decisions:

*Converting data into data_long which long format for flexible plotting multiple countries.

*Used clean,minimal themes to reduce visual clutter.

*included clear axies labels, titles, and legends


---

## Responsible Use of AI (RVE Note)

Complete one of the two options below.

### **If you used AI:**

Include the following block for each interaction.

```
- Tool/model used: ChatGpt
- Prompt (short paraphrase is OK):
  
  Explain the output step by step.
  
- What the AI generated:
   Explained the output for scatter plot, violin plot word to word .
   
- How you verified it:
   after explaination went to chats and crossed checked everything.
- What you changed:
   After understanding i wrote the description in my own words.
   
```

### **If you did NOT use AI:**

> *“I did not use AI for code or interpretation. All design decisions and commentary are my own.”*

---

## Appendix (Optional)

You may place additional plots, tests, or reproducibility info here.

```{r}
# Optional extras
# sessionInfo()
```

---

# **END OF FILE**

---